---
title: "Gun deaths"
author: "Your name"Christina Feng (cef229)
date: today
format: 
  html
    toc: true
    toc-depth:2
    toc-location: left
    theme:
      light: flatly
      dark: darkly
execute:
  echo: false
  cashe: true

---

```{r}
#| label: setup
#| include: false

library(tidyverse)
library(rcis)

data("gun_deaths")
```

```{r}
#| label: youth

youth <- gun_deaths |>
  filter(age <= 65)
```

# Gun deaths by age

We have data about `r nrow(gun_deaths)` individuals killed by guns. Only `r nrow(gun_deaths) - nrow(youth)` are older than 65. The distribution of the remainder is shown below:

```{r}
#| label: youth-dist
#| echo: false
#| fig-width: 8
#| fig- height: 6
#| cashe: true

ggplot(data = youth, mapping = aes(x = age)) + 
  geom_freqpoly(binwidth = 1)
```
This distribution reveals a significant peak in gun deaths at around 20 years of age. This relative peak slowly declines until a small relative peak around 40 years of age. And then the distribution decreases past 50 years old.

# Gun deaths by race

```{r}
#| label: race-dist
#| echo: false
#| fig-width: 8
#| fig-height: 6
#| cache: true


youth |>
  mutate(race = fct_infreq(race) |> fct_rev()) |>
  ggplot(mapping = aes(y = race)) +
  geom_bar() +
  labs(y = "Victim race")
```
