---
title: "Predicting children in hotel bookings"
author: Your name
format: html
---

```{r}
#| label: setup
#| include: false
library(tidyverse)
library(tidymodels)

hotels <- read_csv("data/hotels-lite.csv") |>
  mutate(across(where(is.character), as.factor))
hotels
```

# Your Turn 1

Run the chunk below and look at the output. Then, copy/paste the code and edit to create:

+ a decision tree model for classification 

+ that uses the `C5.0` engine. 

Save it as `tree_mod` and look at the object. What is different about the output?

*Hint: you'll need https://www.tidymodels.org/find/parsnip/*

```{r}
lr_mod <- logistic_reg() |> 
  set_engine(engine = "glm") |> 
  set_mode("classification")
lr_mod
```

```{r}


```

# Your Turn 2

Fill in the blanks. 

Use `initial_split()`, `training()`, and `testing()` to:

1. Split **hotels** into training and test sets. Save the rsplit!

2. Extract the training data and fit your classification tree model.

3. Check the proportions of the `test` variable in each set.

Keep `set.seed(100)` at the start of your code.

*Hint: Be sure to remove every `_` before running the code!*

```{r}
set.seed(100) # Important!
hotels_split <- ________(hotels, prop = 3 / 4)
hotels_train <- ________(hotels_split)
hotels_test <- ________(hotels_split)

# check distribution
count(x = hotels_train, children) |>
  mutate(prop = n / sum(n))
count(x = hotels_test, children) |>
  mutate(prop = n / sum(n))
```

# Your Turn 3

Run the code below. What does it return?

```{r}
set.seed(100)
hotels_folds <- vfold_cv(data = hotels_train, v = 10)
hotels_folds
```

# Your Turn 4

Add a `autoplot()` to visualize the ROC AUC. How well does the model perform?

```{r}
tree_preds <- tree_mod |> 
  fit_resamples(
    children ~ average_daily_rate + stays_in_weekend_nights, 
    resamples = hotels_folds,
    control = control_resamples(save_pred = TRUE)
  )

tree_preds |>
  collect_predictions() |>
  roc_auc(truth = children, .pred_children)

tree_preds |> 
  collect_predictions() |> 
  roc_curve(truth = children, .pred_children) |>
  ________()
```

*Add response here.*

# Acknowledgments

- Materials derived from [Tidymodels, Virtually](https://tmv.netlify.app/site/) by [Allison Hill](https://www.apreshill.com/) and licensed under a [Creative Commons Attribution-ShareAlike 4.0 International (CC BY-SA) License](https://creativecommons.org/licenses/by/4.0/).
- Dataset and some modeling steps derived from [A predictive modeling case study](https://www.tidymodels.org/start/case-study/) and licensed under a [Creative Commons Attribution-ShareAlike 4.0 International (CC BY-SA) License](https://creativecommons.org/licenses/by/4.0/).
